% 2010 California Household Travel Survey
% Linked Trip Summaries
% %%%TIME%%%

```{r setup,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
# Preload needed code chunks
library(knitr)
sub=read_chunk("../R/load-linked-trips.R", labels="llt-sub") 
sub=read_chunk("../R/unweighted-weekday-trips.R", 
               labels="unweighted-wd-trips-sub")
sub=read_chunk("../R/weighted-wd-trips-type.R",
               labels="weighted-wd-trips-sub")
sub=read_chunk("../R/weighted-we-trips-type.R",
               labels="weighted-we-trips-sub")
# temporarily comment this as it is breaking in R 3.1
#sub=read_chunk("../R/weighted-wd-trips-by-region.R",
#               labels="weighted-wd-trips-by-region-sub")
```

# Introduction

These slides present a summary of the method being used to generate linked trip
tables on the 2010 CHTS data.

* A [pdf version](present.pdf) of this document

* The linked trip table is available as a compressed SQL file upon request.

* The code for this project is maintained as a git repository.  Access is
  available upon request.

* PLEASE NOTE: This is a working document very much still a draft. ***It was
  last updated on %%%TIME%%%***

# Change Log

```{.changelog include="../CHANGELOG"}
```

# Target Tools

The following tools have been chosen as the target tools because they are mature
and actively maintained free and open source software.

* Relational datastore: [PostgresSQL](http://www.postgresql.org/)
* Statistical package: [R](http://www.r-project.org)
* Source code management: [git](http://gitscm.org)

Optional tools include:

* Statistical front end: [Rstudio](http://www.rstudio.com/)

# Setting up your system

## Overview of files in this project

This project is organized as follows

```{.sh}
present/               # documentation and reporting
present/present.Rmd    #   This Rmarkdown file
R/                     # R scripts, used by the documents in present
sql/                   # SQL processing scripts
sql/chts-2010.zip      #   zipped SQL file that will create the CHTS database in postgresql
```

## Getting up and running quickly

### Install postgresql

1. Download the latest version version of
   [postgresql](http://www.postgresql.org/) for your operating system
   
2. Follow the installation instructions, which should include creating a
   database user and password
   
3. On windows, say yes to the use of `stackbuilder` (see below)
   
### Install PostGIS (optional)

PostGIS provides GIS functionality for postgresql databases.
Installation instructions vary depending on the operating system.  

    If you are using the windows installer, the `stackbuilder`
    installation process will install PostGIS for you.  When you get to
    the "select the applications you would like to install", open
    `Categories=>Spatial Extensions` and select the PostGIS bundle that
    is appropriate for your platform (e.g., 32 vs 64bit)
   
### Install R

To install R, find the installation instructions for you operating
system [here](http://www.r-project.org/).  There is a windows installer.
You should be able to accept the default installation options.


### Install Rstudio (optional, but recommended)

[R studio](http://www.rstudio.com/) provides a helpful Integrated
Development Environment (IDE) for R.

### Install git (optional, but recommended)

[git](http://gitscm.org) is a well known distributed version control
system.  It is incredibly flexible and will allow you to maintain your
projects without risking losing your prior work.  It is designed to
allow users to share work efficiently.  

If you are installing on Windows, the default installation options
should work fine with Rstudio EXCEPT, you need to selection the option
to use `git` from the windows command line.  Doing this makes `git`
accessible to Rstudio, which integrates with git and wraps its
complexity within a GUI.  There is a nice overview
[here](https://support.rstudio.com/hc/en-us/articles/200532077-Version-Control-with-Git-and-SVN?version=0.98.1062&mode=desktop_


### Check out the chts-analysis repository

This describes the process for getting a copy of the linked trip table
analysis project using `Rstudio` and `git`.

1. Start `Rstudio`

2. Create a new project

    1. File=>New Project
    2. Select `Version Control`
    3. Select `Git`
    4. Enter the following:

       Field                                 Value                                    
       -----------                           ----------------------------------       
       Repository                            `https://github.com/crindt/chts-analysis`
       Project Directory Name                `chts-analysis`                          
       Create Project as a subdirectory of   `~`                                      

       This will create a project in your home directory that contains
       the analysis source.

    5. Open up the Rproject file under the R/ directory.
   
### Load the chts-2010 data into postgresql

The steps for this will depend on your operating system.  However, the
procedure should be similar.




# Conversion steps



## Platform

The following conversion was performed on a mid-level workstation running the
13.04 Ubuntu Linux distribution.  Only open source tools were required.  (Use
down arrow to see details)

## Convert raw MS-Access (.mdb) databases for use PostgreSQL

The following shell script uses the `mdb-tools` package and standard unix tools
to convert complete set of tables in the Access database to tab separated value
files, which are then loaded into `PostgreSQL`.

~~~~{.bash include="../sql/export-to-db.sh"}
~~~~

## Generate static linked trip table consistent with 2001 CHTS datasets

The following SQL command is used to generate `tbllinkedtrip`, containing the
linked trip tables for use in `R`

~~~~{.sql include="../sql/clean.sql"}
~~~~

Function definitions...

~~~~{.sql include="../sql/helper-functions.sql"}
~~~~



# Importing data from PostgreSQL to R
```{r llt-sub,eval=TRUE,echo=TRUE,warning=FALSE,message=FALSE,fig.keep='none',cache=TRUE}
```

# Trips by Type

## A note on trip types

For the purposes of this analysis, each linked trip origin and destination was
classified using the following `SQL` functions.  There is room for discussion
here regarding these definitions.  Arrow down to see trip type distributions.

``` {.sql include="../sql/trip-type-functions.sql"}
```

## **Unweighted** weekday trips by type

```{r unweighted-wd-trips-sub,fig.cap="",eval=TRUE,echo=FALSE, warning=FALSE, message=FALSE,fig.width=10,fig.height=5}
```
```{r unweighted-wd-trips-sub,fig.cap="",eval=FALSE,echo=TRUE, warning=TRUE, message=TRUE}
```

## Weighted weekday trips by type

```{r weighted-wd-trips-sub,fig.cap="",eval=TRUE,echo=FALSE, warning=FALSE, message=FALSE,fig.width=10,fig.height=5}
```
```{r weighted-wd-trips-sub,fig.cap="",eval=FALSE,echo=TRUE, warning=TRUE, message=TRUE}
```


## Weighted weekend person trips by type

```{r weighted-we-trips-sub,fig.cap="",eval=TRUE,echo=FALSE, warning=FALSE, message=FALSE,fig.width=10,fig.height=5}
```
```{r weighted-wd-trips-sub,fig.cap="",eval=FALSE,echo=TRUE, warning=TRUE, message=TRUE}
```

# Weighted weekday trips by region

Temporarily disabled due to compatibility issues with R 3.1

<!--
```{r weighted-wd-trips-by-region-sub,fig.cap="",eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE,fig.width=10,fig.height=5,results="asis"}
```
```{r weighted-wd-trips-by-region-sub,fig.cap="",eval=FALSE,echo=TRUE, warning=TRUE, message=TRUE}
```
-->

# Data cleaning



# More to come...

This is a living document that we are actively improving.  Check back often...





