% 2010 California Household Travel Survey
% Linked Trip Summaries
% %%%TIME%%%

```{r setup,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
# Preload needed code chunks
library(knitr)
sub=read_chunk("../R/load-linked-trips.R", labels="llt-sub") 
sub=read_chunk("../R/unweighted-weekday-trips.R", 
               labels="unweighted-wd-trips-sub")
sub=read_chunk("../R/weighted-wd-trips-type.R",
               labels="weighted-wd-trips-sub")
sub=read_chunk("../R/weighted-we-trips-type.R",
               labels="weighted-we-trips-sub")
sub=read_chunk("../R/weighted-wd-trips-by-region.R",
               labels="weighted-wd-trips-by-region-sub")
```

# Introduction

These slides present a summary of the method being used to generate linked trip
tables on the 2010 CHTS data.

* A [pdf version](present.pdf) of this document

* The linked trip table is available as a compressed SQL file upon request.

* The code for this project is maintained as a git repository.  Access is
  available upon request.

* PLEASE NOTE: This is a working document very much still a draft. ***It was
  last updated on %%%TIME%%%***

# Change Log

```{.changelog include="../../CHANGELOG"}
```

# Target Tools

The following tools have been chosen as the target tools because they are mature
and actively maintained free and open source software.

* Relational datastore: [PostgresSQL](http://www.postgresql.org/)
* Statistical package: [R](http://www.r-project.org)
* Source code management: [git](http://gitscm.org)

Optional tools include:

* Statistical front end: [Rstudio](http://www.rstudio.com/)

# Setting up your system

# Conversion steps

## Platform

The following conversion was performed on a mid-level workstation running the
13.04 Ubuntu Linux distribution.  Only open source tools were required.  (Use
down arrow to see details)

## Convert raw MS-Access (.mdb) databases for use PostgreSQL

The following shell script uses the `mdb-tools` package and standard unix tools
to convert complete set of tables in the Access database to tab separated value
files, which are then loaded into `PostgreSQL`.

~~~~{.bash include="../sql/export-to-db.sh"}
~~~~

## Generate static linked trip table consistent with 2001 CHTS datasets

The following SQL command is used to generate `tbllinkedtrip`, containing the
linked trip tables for use in `R`

~~~~{.sql include="../sql/clean.sql"}
~~~~

Function definitions...

~~~~{.sql include="../sql/helper-functions.sql"}
~~~~



# Importing data from PostgreSQL to R
```{r llt-sub,eval=TRUE,echo=TRUE,warning=FALSE,message=FALSE,fig.keep='none',cache=TRUE}
```

# Trips by Type

## A note on trip types

For the purposes of this analysis, each linked trip origin and destination was
classified using the following `SQL` functions.  There is room for discussion
here regarding these definitions.  Arrow down to see trip type distributions.

``` {.sql include="../sql/trip-type-functions.sql"}
```

## **Unweighted** weekday trips by type

```{r unweighted-wd-trips-sub,fig.cap="",eval=TRUE,echo=FALSE, warning=FALSE, message=FALSE,fig.width=10,fig.height=5}
```
```{r unweighted-wd-trips-sub,fig.cap="",eval=FALSE,echo=TRUE, warning=TRUE, message=TRUE}
```

## Weighted weekday trips by type

```{r weighted-wd-trips-sub,fig.cap="",eval=TRUE,echo=FALSE, warning=FALSE, message=FALSE,fig.width=10,fig.height=5}
```
```{r weighted-wd-trips-sub,fig.cap="",eval=FALSE,echo=TRUE, warning=TRUE, message=TRUE}
```


## Weighted weekend person trips by type

```{r weighted-we-trips-sub,fig.cap="",eval=TRUE,echo=FALSE, warning=FALSE, message=FALSE,fig.width=10,fig.height=5}
```
```{r weighted-wd-trips-sub,fig.cap="",eval=FALSE,echo=TRUE, warning=TRUE, message=TRUE}
```

# Weighted weekday trips by region

```{r weighted-wd-trips-by-region-sub,fig.cap="",eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE,fig.width=10,fig.height=5,results="asis"}
```
```{r weighted-wd-trips-by-region-sub,fig.cap="",eval=FALSE,echo=TRUE, warning=TRUE, message=TRUE}
```

# Data cleaning



# More to come...

This is a living document that we are actively improving.  Check back often...





